---
title: KEGG pathway analysis for control data
author: Aaron Lun
date: 12 September 2016
---

```{r, echo=FALSE, results='hide'}
knitr::opts_chunk$set(error=FALSE, message=FALSE)
knitr::opts_knit$set(width=100)
options(width=100)
```

# Examining DE in gene sets for the control comparisons

Checking for DE in GO- or KEGG-based gene sets.
Setting up a function to do this quickly and efficiently for each contrast.

```{r}
dir.create("pathways", showWarning=FALSE)
dir.create("pathways/results_de", showWarning=FALSE)
dir.create("pathways/results_lfc", showWarning=FALSE)
threshold <- readRDS("threshold_lfc.rds")
require(limma)
require(org.Hs.eg.db)
test_gene_sets <- function(incoming, use.threshold=NA) {
    # Converting names and finding DE genes.
    stuff <- read.table(incoming, header=TRUE, stringsAsFactor=FALSE)
    entrez <- select(org.Hs.eg.db, keys=stuff$ENSEMBL, keytype="ENSEMBL", column="ENTREZID")
    entrez <- entrez[match(stuff$ENSEMBL, entrez$ENSEMBL),]
    if (is.na(use.threshold)) { 
        is.de <- stuff$P.Value <= threshold 
    } else {
        is.de <- stuff$adj.P.Val <= use.threshold
    }
    is.de <- is.de & !is.na(entrez$ENTREZID)

    # Running KEGG.
    kegg <- kegga(entrez$ENTREZID[is.de], species="Hs")
    kegg <- kegg[order(kegg$P.DE),]
    write.table(file=file.path("pathways", dirname(incoming), 
        paste0("kegg_", basename(incoming))), kegg, col.names=NA, 
        sep="\t", quote=FALSE)
       
    # Running GO. 
    go <- goana(entrez$ENTREZID[is.de], species="Hs")
    go <- go[order(go$P.DE),]
    write.table(file=file.path("pathways", dirname(incoming), 
        paste0("go_", basename(incoming))), go, col.names=NA, 
        sep="\t", quote=FALSE)

    # Also reporting the DE genes.
    return(list(kegg=kegg, go=go, 
        de=data.frame(SYMBOL=stuff$SYMBOL[is.de], entrez[is.de,], stringsAsFactors=FALSE)))
}
```

Also maintaining a list of genes involved in terms of interest, e.g., cell cycle, DNA replication.

```{r}
of.interest <- list()
allkegg <- getGeneKEGGLinks(species.KEGG = "hsa", convert = FALSE)
of.interest$CellCycle <- allkegg$GeneID[allkegg$PathwayID=="path:hsa04110"]
of.interest$DNARep <- allkegg$GeneID[allkegg$PathwayID=="path:hsa03030"]
```

## Gene set tests for the siRNA comparisons

First, the Ambion controls against the cells.

```{r}
out <- test_gene_sets("results_de/siRNA_Ambion_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Same again for the log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/siRNA_Ambion_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Next, the Dharmacon controls.

```{r}
out <- test_gene_sets("results_de/siRNA_Dharmacon_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Same again for the log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/siRNA_Dharmacon_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Finally, the two sets of controls against each other.

```{r}
out <- test_gene_sets("results_de/siRNA_Ambion_vs_Dharmacon.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Same again for the log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/siRNA_Ambion_vs_Dharmacon.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

## Gene set tests for the LNA comparisons

First testing the transfection effect.

```{r}
out <- test_gene_sets("results_de/LNA_trans_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Again, with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/LNA_trans_vs_cells.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Testing the effect of negative guide A transfection.

```{r}
out <- test_gene_sets("results_de/LNA_controlA_vs_trans.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Again, with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/LNA_controlA_vs_trans.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Testing the effect of negative guide B transfection.

```{r}
out <- test_gene_sets("results_de/LNA_controlB_vs_trans.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Again, with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/LNA_controlB_vs_trans.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Next, the two controls against each other.

```{r}
out <- test_gene_sets("results_de/LNA_controlA_vs_controlB.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Again, with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/LNA_controlA_vs_controlB.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Finally, as a special treat, we test the effect of 289 itself (we use the supplementary tables in which the DE genes have already been defined, so all genes are included).

```{r}
out <- test_gene_sets("results_lfc/SUPP_combined_LNA_289.txt", use.threshold=1)
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

## Gene set tests for the CRISPRi comparisons

Testing the negative guide-treated cells against the Cas9-only clone 2.

```{r}
out <- test_gene_sets("results_de/CRISPRi_negguide2_vs_clone2_I.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Repeating with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/CRISPRi_negguide2_vs_clone2_I.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

And again, for the Cas9-only clone 2 against WT cells.

```{r}
out <- test_gene_sets("results_de/CRISPRi_clone2_vs_cells_I.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Repeating with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/CRISPRi_clone2_vs_cells_I.txt")
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Finally, for the common set in all clones against WT cells (the threshold can be anything, as only the significant genes have non-unity adjusted p-values in this table).

```{r}
out <- test_gene_sets("results_de/combined_clones.txt", use.threshold=0.05)
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

Repeating with a log-fold change threshold.

```{r}
out <- test_gene_sets("results_lfc/combined_clones.txt", use.threshold=0.05)
head(out$kegg)
head(out$go)
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$CellCycle]
out$de$SYMBOL[out$de$ENTREZID %in% of.interest$DNARep]
```

# Session information

```{r}
sessionInfo()
```
