---
title: Computing the signal-to-noise ratio for DE analyses
author: Aaron Lun
date: 6 March 2018
output:
  html_document:
    toc: true
    toc_float: true
    depth: 3
    number_sections: true
    theme: united
    highlight: tango
    fig_caption: false
---

```{r, echo=FALSE, results='hide'}
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(width=100, fig.path="figure-snr/")
dir.create('figure-snr', showWarning=FALSE)
options(width=100)
```

# Pulling in the data

We pull in the old data (with the controls) and new data (with the comparisons for ch-TOG and MALAT1).

```{r}
library(edgeR)
yold <- readRDS("../../analysis/differential/object.rds")
ynew <- readRDS("object.rds")
```

We make sure that they have the same gene orders.

```{r}
shared <- intersect(rownames(yold), rownames(ynew))
yold <- yold[shared,]
ynew <- ynew[shared,]
```

We augment the annotation.

```{r}
ynew$samples$group <- paste0(sub(" ", "_", ynew$samples$group), ".batch_6")
ynew$samples$block <- paste0("6", ynew$samples$block)
y <- cbind(yold, ynew)
```

To equalize power, we only keep two replicates for each condition.
We also discard conditions that are not relevant to our comparison.

```{r}
useful.groups <- c( # CRISPRi:
    "CRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_1.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.289_guide_9.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.H19_guide_2.batch_5",
    "TOG94.batch_6", "Malat84.batch_6", "NC2_CRISPRi.batch_6",
    # LNA:
    "LNA.wild_type_genotype.Negative_control_A.batch_4",
    "LNA.wild_type_genotype.Negative_control_B.batch_4",
    "LNA.wild_type_genotype.Negative_control_B.batch_1",
    "LNA.wild_type_genotype.289_LNA.batch_1",
    "Malat_LNA.batch_6", "LNA_A.batch_6",
    # RNAi:
    "RNA_interference.wild_type_genotype.Ambion_control.batch_1",
    "RNA_interference.wild_type_genotype.Dharmacon_control.batch_1",
    "RNA_interference.wild_type_genotype.289_siRNA.batch_1",
    "TOGsi.batch_6", "Con_Dharm.batch_6")
keep <- y$samples$group %in% useful.groups
y <- y[,keep]
summary(keep)
useful.blocks <- c("12", "13", "42", "43", "52", "53", "61", "62", "65", "66")
keep <- y$samples$block %in% useful.blocks
y <- y[,keep]
summary(keep)
```

# Setting up the DE analysis

Setting up the design matrix.
Note some unestimable coefficients need to be removed.

```{r}
library(limma)
g <- factor(y$samples$group)
b <- factor(y$samples$block)
design <- model.matrix(~ 0 + g + b)
c(ncol(design), qr(design)$rank)
design <- design[,!colnames(design)%in% c("b42", "b52", "b61", "b65")]
c(ncol(design), qr(design)$rank)
```

We normalize and estimate the variance with our restricted subset of libraries.
We use the same model fit across all our contrasts so that the samew vairance estimate is used.
This avoids issues with differences in power between contrasts.

```{r}
y <- calcNormFactors(y)
v <- voom(y, design, plot=TRUE)
```

Fitting a linear model.

```{r}
fit <- lmFit(v, design)
fit <- treat(lfc=0.5, fit, robust=TRUE)
plotSA(fit)
```

We set up a function to generate ROC curves based on the p-values.
The idea is to define the ratio of the number of DE genes in the knockdown samples to the number of DE genes between the negative controls.

```{r}
GEN <- function(pos, neg) {
    findInterval(sort(neg), sort(pos))
}
```

# RNAi comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design, 
    gRNA_interference.wild_type_genotype.Ambion_control.batch_1 -
    gRNA_interference.wild_type_genotype.Dharmacon_control.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.neg <- fit2$p.value
rnai.neg.lfc <- fit2$coefficients
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design, 
    gRNA_interference.wild_type_genotype.289_siRNA.batch_1 - 
    gRNA_interference.wild_type_genotype.Dharmacon_control.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.289 <- fit2$p.value
rnai.289.lfc <- fit2$coefficients
```

And same for TOG.

```{r}
con <- makeContrasts(levels=design, gTOGsi.batch_6 - gCon_Dharm.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.TOG <- fit2$p.value
rnai.TOG.lfc <- fit2$coefficients
```

We now look at these ROC curves.

```{r}
plot(GEN(rnai.TOG, rnai.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    main="RNAi")
lines(GEN(rnai.289, rnai.neg), col="red", lwd=2)
legend("topleft", col=c("black", "red"), legend=c("ch-TOG", "289"), lwd=2)
```

# LNA comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design, 
    gLNA.wild_type_genotype.Negative_control_A.batch_4 -
    gLNA.wild_type_genotype.Negative_control_B.batch_4)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.neg <- fit2$p.value
lna.neg.lfc <- fit2$coefficients
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design,
    gLNA.wild_type_genotype.Negative_control_B.batch_1 -
    gLNA.wild_type_genotype.289_LNA.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.289 <- fit2$p.value
lna.289.lfc <- fit2$coefficients
```

And same for MALAT1.

```{r}
con <- makeContrasts(levels=design, gMalat_LNA.batch_6 - gLNA_A.batch_6) 
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.MALAT <- fit2$p.value
lna.MALAT.lfc <- fit2$coefficients
```

We now look at these ROC curves.

```{r}
plot(GEN(lna.MALAT, lna.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    col="blue", main="LNA")
lines(GEN(lna.289, lna.neg), col="red", lwd=2)
legend("topleft", col=c("blue", "red"), legend=c("MALAT1", "289"), lwd=2)
```

# CRISPRi comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design,
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_1.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.neg <- fit2$p.value
crispri.neg.lfc <- fit2$coefficients
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design,
    gCRISPRi.heterogeneous_dCas9_KRAB.289_guide_9.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.289 <- fit2$p.value
crispri.289.lfc <- fit2$coefficients
```

And same for H19.

```{r}
con <- makeContrasts(levels=design, 
    gCRISPRi.heterogeneous_dCas9_KRAB.H19_guide_2.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.H19 <- fit2$p.value
crispri.H19.lfc <- fit2$coefficients
```

And same for TOG.

```{r}
con <- makeContrasts(levels=design, gTOG94.batch_6 - gNC2_CRISPRi.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.TOG <- fit2$p.value
crispri.TOG.lfc <- fit2$coefficients
```

And same for MALAT1.

```{r}
con <- makeContrasts(levels=design, gMalat84.batch_6 - gNC2_CRISPRi.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.MALAT <- fit2$p.value
crispri.MALAT.lfc <- fit2$coefficients
```

We now look at these ROC curves.

```{r}
plot(GEN(crispri.MALAT, crispri.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    col="blue", main="CRISPRi")
lines(GEN(crispri.289, crispri.neg), col="red", lwd=2)
lines(GEN(crispri.TOG, crispri.neg), col="black", lwd=2)
lines(GEN(crispri.H19, crispri.neg), col="orange", lwd=2)
legend("topleft", col=c("blue", "red", "black", "orange"), 
    legend=c("MALAT1", "289", "ch-TOG", "H19"), lwd=2)
```

# Putting everything together

We create a three-panel PDF where the ROC curves for all genes are combined, arranged by technology.

```{r}
xlab <- "Number of DE genes between negative controls"
ylab <- "Number of DE genes upon knockdown"
xlim <- c(0, 200)
ylim <- c(0, 2000)
colors <- c(`289`="grey50", H19="blue", MALAT1="red", `ch-TOG`="goldenrod")
xpos <- xlim[2] - 10
pdf("pics/roc.pdf")

# For RNAi
o.rnai.TOG <-GEN(rnai.TOG, rnai.neg)
plot(o.rnai.TOG, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="RNAi")
lines(o.rnai.TOG, lwd=3, col=colors["ch-TOG"])
text(xpos, o.rnai.TOG[xlim[2]], "ch-TOG", pos=3)

o.rnai.289 <- GEN(rnai.289, rnai.neg)
lines(o.rnai.289, lwd=5)
lines(o.rnai.289, col=colors["289"], lwd=3)
text(xpos, o.rnai.289[xlim[2]], "289", pos=3)

# For LNA.
o.lna.MALAT <- GEN(lna.MALAT, lna.neg)
plot(o.lna.MALAT, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="LNA")
lines(o.lna.MALAT, col=colors["MALAT1"], lwd=3) 
text(xpos, o.lna.MALAT[xlim[2]], "MALAT1", pos=3)

o.lna.289 <- GEN(lna.289, lna.neg)
lines(o.lna.289, lwd=5)
lines(o.lna.289, col=colors["289"], lwd=3)
text(xpos, o.lna.289[xlim[2]], "289", pos=3)

# For CRISPRi
o.crispri.MALAT <- GEN(crispri.MALAT, crispri.neg)
plot(o.crispri.MALAT, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="CRISPRi")
lines(o.crispri.MALAT, col=colors["MALAT1"], lwd=3) 
text(xpos, o.crispri.MALAT[xlim[2]], "MALAT1", pos=3)

o.crispri.289 <- GEN(crispri.289, crispri.neg) 
lines(o.crispri.289, lwd=5)
lines(o.crispri.289, col=colors["289"], lwd=3)
text(xpos, o.crispri.289[xlim[2]], "289", pos=3)

o.crispri.TOG <- GEN(crispri.TOG, crispri.neg)
lines(o.crispri.TOG, lwd=5)
lines(o.crispri.TOG, col=colors["ch-TOG"], lwd=3)
text(xpos, o.crispri.TOG[xlim[2]] - 20, "ch-TOG", pos=1)

o.crispri.H19 <- GEN(crispri.H19, crispri.neg)
lines(o.crispri.H19, lwd=5)
lines(o.crispri.H19, col=colors["H19"], lwd=3)
text(xpos, o.crispri.H19[xlim[2]], "H19", pos=1)
dev.off()
```

We also compile the SNRs for the p-values at a value of 200 in the denominator.
This provides us with stable estimates of the SNR that can be reasonably compared between methods. 

```{r}
o.rnai.TOG[200]/200
o.rnai.289[200]/200
o.lna.289[200]/200
o.lna.MALAT[200]/200
o.crispri.TOG[200]/200
o.crispri.289[200]/200
o.crispri.MALAT[200]/200
o.crispri.H19[200]/200
```

We repeat this with the log-fold change, where we are looking for genes with large absolute log-fold changes rather than small p-values.
This is probably less reliable than the p-values, but seems to give roughly similar conclusions.

```{r}
xlab <- "Genes with large log-fold changes between negative controls"
ylab <- "Genes with large log-fold changes upon knockdown"
pdf("pics/roc_lfc.pdf")

# For RNAi
o.rnai.TOG.lfc <- GEN(-abs(rnai.TOG.lfc), -abs(rnai.neg.lfc))
plot(o.rnai.TOG.lfc, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="RNAi")
lines(o.rnai.TOG.lfc, lwd=3, col=colors["ch-TOG"])
text(xpos, o.rnai.TOG.lfc[xlim[2]], "ch-TOG", pos=3)

o.rnai.289.lfc <- GEN(-abs(rnai.289.lfc), -abs(rnai.neg.lfc))
lines(o.rnai.289.lfc, lwd=5)
lines(o.rnai.289.lfc, col=colors["289"], lwd=3)
text(xpos, o.rnai.289.lfc[xlim[2]], "289", pos=3)

# For LNA.
o.lna.MALAT.lfc <- GEN(-abs(lna.MALAT.lfc), -abs(lna.neg.lfc))
plot(o.lna.MALAT.lfc, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="LNA")
lines(o.lna.MALAT.lfc, col=colors["MALAT1"], lwd=3) 
text(xpos, o.lna.MALAT.lfc[xlim[2]], "MALAT1", pos=3)

o.lna.289.lfc <- GEN(-abs(lna.289.lfc), -abs(lna.neg.lfc))
lines(o.lna.289.lfc, lwd=5)
lines(o.lna.289.lfc, col=colors["289"], lwd=3)
text(xpos, o.lna.289.lfc[xlim[2]], "289", pos=3)

# For CRISPRi
o.crispri.MALAT.lfc <- GEN(-abs(crispri.MALAT.lfc), -abs(crispri.neg.lfc))
plot(o.crispri.MALAT.lfc, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="CRISPRi")
lines(o.crispri.MALAT.lfc, col=colors["MALAT1"], lwd=3) 
text(xpos, o.crispri.MALAT.lfc[xlim[2]], "MALAT1", pos=3)

o.crispri.289.lfc <- GEN(-abs(crispri.289.lfc), -abs(crispri.neg.lfc))
lines(o.crispri.289.lfc, lwd=5)
lines(o.crispri.289.lfc, col=colors["289"], lwd=3)
text(xpos, o.crispri.289.lfc[xlim[2]], "289", pos=3)

o.crispri.TOG.lfc <- GEN(-abs(crispri.TOG.lfc), -abs(crispri.neg.lfc))
lines(o.crispri.TOG.lfc, lwd=5)
lines(o.crispri.TOG.lfc, col=colors["ch-TOG"], lwd=3)
text(xpos, o.crispri.TOG.lfc[xlim[2]] - 20, "ch-TOG", pos=1)

o.crispri.H19.lfc <- GEN(-abs(crispri.H19.lfc), -abs(crispri.neg.lfc))
lines(o.crispri.H19.lfc, lwd=5)
lines(o.crispri.H19.lfc, col=colors["H19"], lwd=3)
text(xpos, o.crispri.H19.lfc[xlim[2]], "H19", pos=1)
dev.off()
```


# Session info

```{r}
sessionInfo()
```

