---
title: Computing the signal-to-noise ratio for DE analyses
author: Aaron Lun
date: 6 March 2018
output:
  html_document:
    toc: true
    toc_float: true
    depth: 3
    number_sections: true
    theme: united
    highlight: tango
    fig_caption: false
---

```{r, echo=FALSE, results='hide'}
knitr::opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(width=100, fig.path="figure-snr/")
dir.create('figure-snr', showWarning=FALSE)
options(width=100)
```

# Pulling in the data

We pull in the old data (with the controls) and new data (with the comparisons for ch-TOG and MALAT1).

```{r}
library(edgeR)
yold <- readRDS("../../analysis/differential/object.rds")
ynew <- readRDS("object.rds")
```

We make sure that they have the same gene orders.

```{r}
shared <- intersect(rownames(yold), rownames(ynew))
yold <- yold[shared,]
ynew <- ynew[shared,]
```

We augment the annotation.

```{r}
ynew$samples$group <- paste0(sub(" ", "_", ynew$samples$group), ".batch_6")
ynew$samples$block <- paste0("6", ynew$samples$block)
y <- cbind(yold, ynew)
```

To equalize power, we only keep two of each condition.
We also discard conditions that are not relevant to our comparison.

```{r}
useful.groups <- c( # CRISPRi:
    "CRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_1.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.289_guide_9.batch_5",
    "CRISPRi.heterogeneous_dCas9_KRAB.H19_guide_2.batch_5",
    "TOG94.batch_6", "Malat84.batch_6", "NC2_CRISPRi.batch_6",
    # LNA:
    "LNA.wild_type_genotype.Negative_control_A.batch_4",
    "LNA.wild_type_genotype.Negative_control_B.batch_4",
    "LNA.wild_type_genotype.Negative_control_B.batch_1",
    "LNA.wild_type_genotype.289_LNA.batch_1",
    "Malat_LNA.batch_6", "LNA_A.batch_6",
    # RNAi:
    "RNA_interference.wild_type_genotype.Ambion_control.batch_1",
    "RNA_interference.wild_type_genotype.Dharmacon_control.batch_1",
    "RNA_interference.wild_type_genotype.289_siRNA.batch_1",
    "TOGsi.batch_6", "Con_Dharm.batch_6")
keep <- y$samples$group %in% useful.groups
y <- y[,keep]
summary(keep)
useful.blocks <- c("12", "13", "42", "43", "52", "53", "61", "62")
keep <- y$samples$block %in% useful.blocks
y <- y[,keep]
summary(keep)
```

# Setting up the DE analysis

Setting up the design matrix.
Note some unestimable coefficients need to be removed.

```{r}
library(limma)
g <- factor(y$samples$group)
b <- factor(y$samples$block)
design <- model.matrix(~ 0 + g + b)
c(ncol(design), qr(design)$rank)
design <- design[,!colnames(design)%in% c("b42", "b52", "b61")]
c(ncol(design), qr(design)$rank)
```

We normalize and estimate the variance with our restricted subset of libraries, to allow us to compare p-values between contrasts.

```{r}
y <- calcNormFactors(y)
v <- voom(y, design, plot=TRUE)
```

Fitting a linear model.

```{r}
fit <- lmFit(v, design)
fit <- treat(lfc=0.5, fit, robust=TRUE)
plotSA(fit)
```

We set up a function to generate ROC curves.

```{r}
GEN <- function(pos, neg) {
    findInterval(sort(neg), sort(pos))
}
```

# RNAi comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design, 
    gRNA_interference.wild_type_genotype.Ambion_control.batch_1 -
    gRNA_interference.wild_type_genotype.Dharmacon_control.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.neg <- fit2$p.value
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design, 
    gRNA_interference.wild_type_genotype.289_siRNA.batch_1 - 
    gRNA_interference.wild_type_genotype.Dharmacon_control.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.289 <- fit2$p.value
```

And same for TOG.

```{r}
con <- makeContrasts(levels=design, gTOGsi.batch_6 - gCon_Dharm.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
rnai.TOG <- fit2$p.value
```

We now look at these ROC curves.

```{r}
plot(GEN(rnai.TOG, rnai.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    main="RNAi")
lines(GEN(rnai.289, rnai.neg), col="red", lwd=2)
legend("topleft", col=c("black", "red"), legend=c("ch-TOG", "289"), lwd=2)
```

# LNA comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design, 
    gLNA.wild_type_genotype.Negative_control_A.batch_4 -
    gLNA.wild_type_genotype.Negative_control_B.batch_4)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.neg <- fit2$p.value
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design,
    gLNA.wild_type_genotype.Negative_control_B.batch_1 -
    gLNA.wild_type_genotype.289_LNA.batch_1)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.289 <- fit2$p.value
```

And same for MALAT1.

```{r}
con <- makeContrasts(levels=design, gMalat_LNA.batch_6 - gLNA_A.batch_6) 
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
lna.MALAT <- fit2$p.value
```

We now look at these ROC curves.

```{r}
plot(GEN(lna.MALAT, lna.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    col="blue", main="LNA")
lines(GEN(lna.289, lna.neg), col="red", lwd=2)
legend("topleft", col=c("blue", "red"), legend=c("MALAT1", "289"), lwd=2)
```

# CRISPRi comparisons

First we compare between negative controls.

```{r}
con <- makeContrasts(levels=design,
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_1.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.neg <- fit2$p.value
```

We then compare 289 knockdown.

```{r}
con <- makeContrasts(levels=design,
    gCRISPRi.heterogeneous_dCas9_KRAB.289_guide_9.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.289 <- fit2$p.value
```

And same for H19.

```{r}
con <- makeContrasts(levels=design, 
    gCRISPRi.heterogeneous_dCas9_KRAB.H19_guide_2.batch_5 - 
    gCRISPRi.heterogeneous_dCas9_KRAB.Negative_control_guide_2.batch_5)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.H19 <- fit2$p.value
```

And same for TOG.

```{r}
con <- makeContrasts(levels=design, gTOG94.batch_6 - gNC2_CRISPRi.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.TOG <- fit2$p.value
```

And same for MALAT1.

```{r}
con <- makeContrasts(levels=design, gMalat84.batch_6 - gNC2_CRISPRi.batch_6)
fit2 <- contrasts.fit(fit, con)
fit2 <- treat(lfc=0.5, fit2, robust=TRUE)
crispri.MALAT <- fit2$p.value
```

We now look at these ROC curves.

```{r}
plot(GEN(crispri.MALAT, crispri.neg), xlim=c(0, 200), ylim=c(0, 2000), type="l", lwd=2,
    xlab="Number of false positives", ylab="Number of DE genes upon knockdown",
    col="blue", main="CRISPRi")
lines(GEN(crispri.289, crispri.neg), col="red", lwd=2)
lines(GEN(crispri.TOG, crispri.neg), col="black", lwd=2)
lines(GEN(crispri.H19, crispri.neg), col="orange", lwd=2)
legend("topleft", col=c("blue", "red", "black", "orange"), 
    legend=c("MALAT1", "289", "ch-TOG", "H19"), lwd=2)
```

# Putting everything together

We create a three-panel PDF where everything is stuck together, arranged by technology.

```{r}
xlab <- "Number of DE genes between negative controls"
ylab <- "Number of DE genes upon knockdown"
xlim <- c(0, 200)
ylim <- c(0, 2000)

colors <- c(`289`="grey50", H19="blue", MALAT1="red", `ch-TOG`="goldenrod")
xpos <- xlim[2] - 10

pdf("pics/roc.pdf")

# For RNAi
o.rnai.TOG <-GEN(rnai.TOG, rnai.neg)
plot(o.rnai.TOG, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="RNAi")
lines(o.rnai.TOG, lwd=3, col=colors["ch-TOG"])
text(xpos, o.rnai.TOG[xlim[2]], "ch-TOG", pos=3)

o.rnai.289 <- GEN(rnai.289, rnai.neg)
lines(o.rnai.289, lwd=5)
lines(o.rnai.289, col=colors["289"], lwd=3)
text(xpos, o.rnai.289[xlim[2]], "289", pos=3)

# For LNA.
o.lna.MALAT <- GEN(lna.MALAT, lna.neg)
plot(o.lna.MALAT, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="LNA")
lines(o.lna.MALAT, col=colors["MALAT1"], lwd=3) 
text(xpos, o.lna.MALAT[xlim[2]], "MALAT1", pos=3)

o.lna.289 <- GEN(lna.289, lna.neg)
lines(o.lna.289, lwd=5)
lines(o.lna.289, col=colors["289"], lwd=3)
text(xpos, o.lna.289[xlim[2]], "289", pos=3)

# For CRISPRi
o.crispri.MALAT <- GEN(crispri.MALAT, crispri.neg)
plot(o.crispri.MALAT, xlim=xlim, ylim=ylim, type="l", lwd=5,
    xlab=xlab, ylab=ylab, main="CRISPRi")
lines(o.crispri.MALAT, col=colors["MALAT1"], lwd=3) 
text(xpos, o.crispri.MALAT[xlim[2]], "MALAT1", pos=3)

o.crispri.289 <- GEN(crispri.289, crispri.neg) 
lines(o.crispri.289, lwd=5)
lines(o.crispri.289, col=colors["289"], lwd=3)
text(xpos, o.crispri.289[xlim[2]], "289", pos=3)

o.crispri.TOG <- GEN(crispri.TOG, crispri.neg)
lines(o.crispri.TOG, lwd=5)
lines(o.crispri.TOG, col=colors["ch-TOG"], lwd=3)
text(xpos, o.crispri.TOG[xlim[2]] - 20, "ch-TOG", pos=1)

o.crispri.H19 <- GEN(crispri.H19, crispri.neg)
lines(o.crispri.H19, lwd=5)
lines(o.crispri.H19, col=colors["H19"], lwd=3)
text(xpos, o.crispri.H19[xlim[2]], "H19", pos=1)
dev.off()
```

# Session info

```{r}
sessionInfo()
```

